---
title: "Merge meterological and respiration data"
author: "K Todd-Brown (ktoddbrown@gmail.com)"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
---

```{r setUp, message=FALSE}
library(tidyverse)
library(knitr)

sessionInfo()
```

#Load data
##Meterological
```{r loadMet}

load('temp/MetData20170903.RData')
met.df <- data_addWetL.df

ggplot(met.df, aes(x=Air.temperature, y=Relative.Humidity)) +
  geom_hex() +
  facet_wrap(~Location)
```

##Cumulative hours in wet events
```{r cumWetHours}
#Taken from MetData.Rmd
# compareTime.df <- data_addWetL.df %>%
#   ungroup() %>% group_by(Location) %>%
#   summarize(total.hr=sum(duration.hr, na.rm=TRUE),
#          rain.hr=sum(duration.hr*(Rain > 0), na.rm=TRUE),
#          highRH.hr=sum(duration.hr*(Relative.Humidity>80), na.rm=TRUE),
#          wet.hr=sum(duration.hr*(grepl('wet', Wet.factor)), na.rm=TRUE),
#          wet_by_RH.hr=sum(duration.hr*percentWet_byRH, na.rm=TRUE))

cumWet.df <- met.df %>% 
  arrange(Location, Date) %>%
  ungroup() %>% group_by(Location)  %>%
  mutate(duration.hr = if_else(is.na(duration.hr), 0, duration.hr)) %>%
  mutate(rain.cumHr=cumsum(duration.hr*(is.finite(Rain) & Rain > 0)),
         highRH.cumHr=cumsum(duration.hr*(Relative.Humidity>80)),
         wet.cumHr=cumsum(duration.hr*(grepl('wet', Wet.factor))),
         wet_by_RH.cumHr=cumsum(duration.hr*percentWet_byRH)) %>%
  select(Location, Date, duration.hr, rain.cumHr, highRH.cumHr, wet.cumHr, wet_by_RH.cumHr) %>%
  gather(key=wetDef, value=`cum. hours`, 
         rain.cumHr, highRH.cumHr, wet.cumHr, wet_by_RH.cumHr) %>%
  mutate(`Wet Definition` = recode(wetDef, rain.cumHr='Rain', highRH.cumHr="RH > 80%", 
         wet.cumHr="Wetness sensor", wet_by_RH.cumHr="Wet(RH) probability"))

ggplot(cumWet.df) +
  geom_line(aes(x=Date, y=`cum. hours`, color=`Wet Definition`)) +
  facet_wrap(~Location, ncol=1)
```

##Respiration
```{r loadResp}
load('temp/RespData.RData')

resp.df <- data.df %>%
  filter(sterile_non=='nonsterile')

ggplot(resp.df) + geom_histogram(aes(x=Flux_g.C_glitter.1_d.1))
ggplot(resp.df) + geom_point(aes(x=GravMoist, y=Flux_g.C_glitter.1_d.1)) +
  geom_vline(xintercept=c(5, 25))
```

#Extrapolate respiration using met data
Since we can't see a pattern with temperature or moisture (see RespDataProcessing).
Since we have the wetness sensor data we are using that to bin dry, slightly wet, and wet conditions.
We are seperating events triggered by rain vs non-rainfall via wet slopes.

```{r summarizeResp}
resp.short <- resp.df %>%
  filter(is.finite(Flux_g.C_glitter.1_d.1)) %>%
  select(Site_FEDGbbIowa, GravMoist, Flux_g.C_glitter.1_d.1) %>%
  mutate(Wet.factor = factor(ifelse(GravMoist < 5, 'dry', 
                             ifelse(GravMoist < 25, 'slightly.wet', 'wet')),
                             c('dry', 'slightly.wet', 'wet'))) %>%
  group_by(Wet.factor) %>%
  summarize(Flux000 = quantile(Flux_g.C_glitter.1_d.1, 0),
            Flux005 = quantile(Flux_g.C_glitter.1_d.1, 0.05),
            Flux025 = quantile(Flux_g.C_glitter.1_d.1, 0.25),
            Flux050 = quantile(Flux_g.C_glitter.1_d.1, 0.50),
            Flux075 = quantile(Flux_g.C_glitter.1_d.1, 0.75),
            Flux095 = quantile(Flux_g.C_glitter.1_d.1, 0.95),
            Flux100 = quantile(Flux_g.C_glitter.1_d.1, 1)) %>%
  mutate(units='g.C_glitter.1_d.1')

print(select(resp.short, Wet.factor, units, Flux000, Flux050, Flux100))

```

#Calculate resipration range from different definnitions of wet time
```{r compare}
wet_respRate <- resp.short %>% filter(Wet.factor == 'wet') %>% select(-Wet.factor, -units)

compareResp.df <- compareTime.df %>% 
  tidyr::gather(Type, value, -Location, -total.hr) %>% 
  filter(value > 0) %>% 
  mutate(Flux000_gC_gLitter.1 = wet_respRate$Flux000*value/24,
         Flux005_gC_gLitter.1 = wet_respRate$Flux005*value/24,
         Flux025_gC_gLitter.1 = wet_respRate$Flux025*value/24,
         Flux050_gC_gLitter.1 = wet_respRate$Flux050*value/24,
         Flux075_gC_gLitter.1 = wet_respRate$Flux075*value/24,
         Flux095_gC_gLitter.1 = wet_respRate$Flux095*value/24,
         Flux100_gC_gLitter.1 = wet_respRate$Flux100*value/24,
         Location=recode_factor(Location, FED="Namib East", 
                                GBB="Namib West", IA="Iowa", SEV="Sevilleta"),
         `Wet Definition`=recode_factor(Type, rain.hr = "Rain",
                                 highRH.hr="RH > 80%",
                                 wet.hr="Wetness sensor",
                                 wet_by_RH.hr ='Wet(RH) probability'))

ggplot(compareResp.df, aes(x=Location, y=value/total.hr, color=`Wet Definition`)) + 
  geom_point() + geom_line() +
  labs(x='Location', y='Fraction of total time by hour', 
       title='Fraction time wet by different definitions') +
  scale_y_log10()

ggplot(compareResp.df, aes(x=`Wet Definition`, #y=Flux050_gC_gLitter.1,
                           lower=Flux025_gC_gLitter.1, 
                           upper=Flux075_gC_gLitter.1, middle= Flux050_gC_gLitter.1,
                           ymin= Flux005_gC_gLitter.1, ymax=Flux095_gC_gLitter.1)) +
  geom_boxplot(stat='identity') + 
  labs(y='C lost over total time period [mass-C %]', x='Wet definition') +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_wrap(~Location, scales='free_y')

#### SE work 18 Jan 2018: Rearranging data to create one simpler graph for main text, and one showing all definitions for supplemental. 

#See NOTE/request for KTB on y-axis units

#New figures Jan 18 2018:

#1) Plot to just look at including NRM+Rain and Rain (Wetness probability), for main text Fig 5
ggplot(subset(compareResp.df, Type %in% c('rain.hr','wet_by_RH.hr')), 
       aes(x=`Wet Definition`, #y=Flux050_gC_gLitter.1,
                           lower=Flux025_gC_gLitter.1, 
                           upper=Flux075_gC_gLitter.1, middle= Flux050_gC_gLitter.1,
                           ymin= Flux005_gC_gLitter.1, ymax=Flux095_gC_gLitter.1)) +
  geom_boxplot(stat='identity') + 
  #scale_x_discrete(breaks=c("rain.hr", 'wet_by_RH.hr'),
  #                 labels=c("Rain", 'Rain + NRM')) +
  labs(y='C lost over total time period [mass-C %]', x='Precipitation type included') +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_wrap(~Location, scales='free_y')

#2) Plot definitions, exclude rain: Wetness sensor, Wetness probability and RH, to show that these estimates all align pretty well (Supplemental)
ggplot(subset(compareResp.df, Type %in% c("highRH.hr", 'wet.hr', 'wet_by_RH.hr')), 
       aes(x=`Wet Definition`, #y=Flux050_gC_gLitter.1,
                           lower=Flux025_gC_gLitter.1, 
                           upper=Flux075_gC_gLitter.1, middle= Flux050_gC_gLitter.1,
                           ymin= Flux005_gC_gLitter.1, ymax=Flux095_gC_gLitter.1)) +
  geom_boxplot(stat='identity') + 
  labs(y='C lost over total time period [mass-C %]', x='Wet definition for NRM') +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_wrap(~Location, scales='free_y')



```
