---
title: "Analyse in meterological data"
author: "K Todd-Brown (ktoddbrown@gmail.com)"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  pdf_document: 
    toc: yes
  html_document: 
    toc: yes
---

This document is to develope the correlation between meterological variables and leaf wetness to seperate moisture events driven by 1) rain, 2) fog, or 3) non-rainfall moisture.

#Set up
```{r setUp, message=FALSE}
library(readxl) #deal with excel data directly
#library(plyr) #aggregate data
library(ggplot2) #make plots
library(reshape2) #cast data from short to long format
library(knitr) #make pretty tables
library(lubridate)
library(GGally)
library(dplyr)


sourceFiles.ls <- list('R/defineSlope.R', 'R/defineEventIndex.R')
temp <- lapply(sourceFiles.ls, source)

sessionInfo()

## Test
```


#Read in GBB data
All times are read in at GMT to preserve local diurnal patterns.

```{r GBB}
metDataFile <- 'data/Gbb/FN GB hour 20150615 20160615.xlsx' 

##Read in units at the start of the table
GBB.meta <- readxl::read_excel(path=metDataFile, sheet="GB", n_max=3)
GBB.meta <- as.data.frame(t(GBB.meta[,-1]))
names(GBB.meta) <- c('variable', 'unit', 'height')
GBB.meta$name <- gsub(' ', '.', row.names(GBB.meta))
print(GBB.meta)

##Read in data
GBB.df <- readxl::read_excel(path=metDataFile, sheet="GB", skip=4, col_names=c('Date', GBB.meta$name),
                             col_types=c('date', rep('numeric', 7)), na=c("NAN", "INF")) %>%
  mutate(Date = as.POSIXct(round.POSIXt(Date, units='mins'), tz='GMT'), 
         Location = 'GBB',
         CumRain = cumsum(Rain)) %>% 
  arrange(Date) %>%
  mutate(duration.hr = c(NA, as.numeric(Date[-1] - Date[-length(Date)], units='hours')))

#Set all GBB 'dry' values to 7999
GBB.df$Leaf.wetness[abs(GBB.df$Leaf.wetness) >= 7999 | 
                      is.na(GBB.df$Leaf.wetness) |
                      GBB.df$Leaf.wetness == -999] <- 7999

rm(GBB.meta)
```

#Read in FED
```{r FED}
FED.df <- read.csv('data/FED/Far_East_met_stat.csv', skip=1, 
                   col.names= c('ID', 'DateStr', 'Air.temperature.globe', 
                    "Wind.direction", 'Wind.speed', 'Gust.speed',
                    'Air.temperature', 'Relative.Humidity')) %>%
  mutate(Date = as.POSIXct(strptime(as.character(DateStr), 
                                    format='%m/%d/%y %I:%M:%S %p', tz='GMT'))) %>%
  select(-ID)



FED2.df <- read.csv('data/FED/Far_East_20161003.csv', header=TRUE, skip=1, stringsAsFactors=FALSE,
                    col.names=  c('ID', 'DateStr', 'Time', 'Air.temperature.globe', 
                    "Wind.direction", 'Wind.speed', 'Gust.speed',
                    'Air.temperature', 'Relative.Humidity'))

#This file switches between DD/MM/YY and MM/DD/YY several times, not worth sorting hrough
startDate <- as.POSIXct(strptime(FED2.df$DateStr[1], format='%d/%m/%y'))
FED2.df$dayOff <- c(0, cumsum(FED2.df$DateStr[-nrow(FED2.df)] != FED2.df$DateStr[-1]))
FED2.df$DDMMYY <- format(startDate + days(FED2.df$dayOff), format='%d/%m/%y')
FED2.df$Date <- as.POSIXct(strptime(sprintf('%s %s', FED2.df$DDMMYY, FED2.df$Time), 
                      format='%d/%m/%y %H:%M:%S'), tz='GMT')#-hours(1)) ##file at GMT+1 to match previous file

FED.df <- rbind(FED.df, FED2.df[, names(FED.df)])
rm(FED2.df)
FED.df <- FED.df[order(FED.df$Date),]
FED.df$MetTimeStamp <- FED.df$Date
FED.df$Date <- as.POSIXct(round(FED.df$Date, units='hour'))
#print(range(round(FED.df$Date)))

##Leaf wetness sensor
FEDlws.df <- read.csv('data/FED/Far_East_LWS.csv', skip=1, 
                      col.names=c('ID', 'Date', 'Wetness.perc')) %>%
  mutate(Date = as.POSIXct(strptime(as.character(Date), format='%m/%d/%y %I:%M:%S %p'), tz='GMT')) %>%
  select(-ID)

#file at GMT+1 but reading in as GMT
FEDlws2.df <- read.csv('data/FED/Far_East_LWS_20161003.csv', skip=1,
                       col.names=c('ID', 'DateStr', 'Time', 'Wetness.perc', sprintf('X%d', 1:7)),
                       stringsAsFactors=FALSE)[,1:4]

#This file switches between DD/MM/YY and MM/DD/YY several times, not worth sorting hrough
startDate <- as.POSIXct(strptime(FEDlws2.df$DateStr[1], format='%d/%m/%y'))
FEDlws2.df$dayOff <- c(0, 
                       cumsum(as.numeric(FEDlws2.df$DateStr[-nrow(FEDlws2.df)] != FEDlws2.df$DateStr[-1]) +
                                 #we are missing July 7 2016 from the dataset
                              as.numeric('8/7/16' == FEDlws2.df$DateStr[-nrow(FEDlws2.df)] &
                                       FEDlws2.df$DateStr[-nrow(FEDlws2.df)] != FEDlws2.df$DateStr[-1])))
FEDlws2.df$DDMMYY <- format(startDate + days(FEDlws2.df$dayOff), format='%d/%m/%y')
FEDlws2.df$Date <- as.POSIXct(strptime(sprintf('%s %s', FEDlws2.df$DDMMYY, FEDlws2.df$Time), 
                format='%d/%m/%y %H:%M:%S'), tz='GMT')

FEDlws.df <- rbind(FEDlws.df, FEDlws2.df[, names(FEDlws.df)])
rm(FEDlws2.df)
FEDlws.df <- FEDlws.df[order(FEDlws.df$Date),]

if(any(unique(minute(FED.df$Date)) != 0)) stop('FED Date assumption not met!')

FEDwet.df <- FEDlws.df %>%
  mutate(shortDate = as.POSIXct(round(Date, units='hours')),
         offHour = as.numeric(abs(Date-as.POSIXct(round(Date, units='hours'))), units='hours')) %>%
  group_by(shortDate) %>%
  summarize(bestTime = Date[which.min(offHour)],
            selectMinute = offHour[which.min(offHour)])

FEDwet.df <- FEDlws.df %>%
  filter(Date %in% FEDwet.df$bestTime) %>%
  mutate(RoundDate = as.POSIXct(round(Date, units='hours'))) %>%
  dplyr::rename(WetnessSensorTimeStamp=Date) %>%
  dplyr::rename(Date=RoundDate) %>% unique

##Precip sensor
##file at GMT+1
FEDrain.df <- read.csv('data/FED/Far_East precip.csv', skip=1, 
                         col.names= c('ID', 'Date', 'CumRain', 'Coupler.detached', 'Host.connected',
                                      'Coupler.attached', 'EOF')) %>%
  mutate(Date = as.POSIXct(strptime(as.character(Date), format='%m/%d/%y %I:%M:%S %p'), tz='GMT')) %>%
  select(Date, CumRain) %>%
  mutate(nearestHour = as.POSIXct(round(Date, units='hour')) )%>%
  group_by(nearestHour) %>%
  summarize(BestTimeStamp=as.POSIXct(Date[which.min(abs(Date-nearestHour))]),
            BestCumRain=CumRain[which.min(abs(Date-nearestHour))]) %>%
  mutate(Rain = c(0, BestCumRain[-1] - BestCumRain[-length(BestCumRain)])) %>%
  dplyr::rename(CumRain=BestCumRain, Date=nearestHour, RainTimeStamp=BestTimeStamp)

FED.df <- FED.df %>%
  merge(FEDwet.df, all=TRUE, by='Date') %>%
  merge(FEDrain.df, all=TRUE, by='Date') %>%
  mutate(Location = 'FED') %>%
  arrange(Date) %>%
  select(-Air.temperature.globe, -Gust.speed) %>%
  mutate(duration.hr = c(NA, as.numeric(Date[-1] - Date[-length(Date)], units='hours')))

FED.df$CumRain[is.na(FED.df$CumRain)] <- 0

rm(FEDlws.df, FEDwet.df, FEDrain.df)
```

#Read in IA
```{r IA}

##2015 isn't covered in the precip file so let's not read in that information
# IA2015.df <- read.xlsx2(file='data/IA/1901_Summer_Prairie_Fall 2015.xls', sheetIndex=1, startRow=2, colClasses=c('numeric', 'character', 'character', rep('numeric', 6)))
# #Personal communciation (Sarah Evans), column names follow same pattern as 2016 
# names(IA2015.df) <- c('ID', 'Date', 'Time', 'Air.temperature.65cm', 'Relative.Humidity.65cm', 'Wetness.perc.65cm', 'Air.temperature', 'Relative.Humidity', 'Wetness.perc')
# #Time in GMT-5 or EST according to header
# IA2015.df$Date <- as.POSIXct(strptime(paste(IA2015.df$Date, IA2015.df$Time), format='%m/%d/%y %H:%M:%S', tz='ETC/GMT-5'))
# IA2015.df$ID <- NULL
# IA2015.df$Time <- NULL

IA2016.df <- readxl::read_excel(path='data/IA/1901_Summer_Prairie_2016.xls', 
                        sheet=1, skip=2,
                        col_names= c('ID', 'DateStr', 'TimeStr', 'Air.temperature.65cm', 'Relative.Humidity.65cm',
                      'Wetness.perc.65cm', 'Air.temperature', 'Relative.Humidity',
                      'Wetness.perc'),
                        col_types='text') %>%
  mutate_at(vars(one_of(c('ID', 'Air.temperature.65cm', 'Relative.Humidity.65cm',
                      'Wetness.perc.65cm', 'Air.temperature', 'Relative.Humidity',
                      'Wetness.perc'))), as.numeric) %>% #Cast as numeric 'cause reading in numeric throws errors
  mutate(MetTimeStamp = as.POSIXct(strptime(paste(DateStr, TimeStr), 
                                      format='%m/%d/%y %H:%M:%S'), tz='GMT')) %>%
  select(-ID)  %>%
  mutate(Date = as.POSIXct(round(MetTimeStamp, units='hour'))) %>%
  group_by(Date) %>%
  summarize(AT=Air.temperature[which.min(abs(MetTimeStamp-Date))],
            RH=Relative.Humidity[which.min(abs(MetTimeStamp-Date))],
            WP = Wetness.perc[which.min(abs(MetTimeStamp-Date))], 
            MD = MetTimeStamp[which.min(abs(MetTimeStamp-Date))]) %>%
  dplyr::rename(Air.temperature=AT, Relative.Humidity=RH, Wetness.perc=WP, MetTimeStamp=MD)

IAprecip.df <- readxl::read_excel(path='data/IA/1901_Summer_Precip_2016.xls', sheet=1, skip=2,
                                  col_names=c('ID', 'DateStr', 'TimeStr', 'Temperature', 'Ext.line.event',
                        'Host.connected', 'Coupler.detached',	'Coupler.attached', 'EOF'),
                          col_types='text') %>%
   mutate_at(vars(one_of(c('ID', 'Temperature'))), as.numeric) %>% #Cast as numeric, reading in numeric errors
        mutate(Date = as.POSIXct(strptime(paste(DateStr, TimeStr), 
                                          format='%m/%d/%y %H:%M:%S'), tz='GMT'), 
#Column E is “events.” Each event is a tip of the “tipping bucket” mechanism inside the rain gauge, and each tip represents 0.2 mm of precip. The “events” accumulate through the record. To measure precip associated with a day or individual storm you simply count the accumulated “tips” and multiply by 0.2 to obtain the total precip in mm.
              CumRain = as.numeric(gsub(',', '', as.character(Ext.line.event)))*0.2) %>%
  filter(is.finite(CumRain)) %>%
  mutate(nearestHour = as.POSIXct(round(Date, units='hour'))) %>%
  group_by(nearestHour) %>%
  summarize(BestTimeStamp=as.POSIXct(Date[which.min(abs(Date-nearestHour))]),
            BestCumRain=CumRain[which.min(abs(Date-nearestHour))]) %>%
  mutate(Rain = c(0, BestCumRain[-1] - BestCumRain[-length(BestCumRain)])) %>%
  dplyr::rename(CumRain=BestCumRain, Date=nearestHour, RainTimeStamp=BestTimeStamp)

#IA.df <- rbind.fill(IA2015.df, IA2016.df)
IA.df <- IA2016.df %>%
  merge(IAprecip.df, all=TRUE, by='Date') %>%
  mutate(Location='IA') %>% 
  arrange(Date) %>%
  mutate(duration.hr = c(NA, as.numeric(Date[-1] - Date[-length(Date)], units='hours')))

IA.df$Rain[is.na(IA.df$Rain)] <- 0
IA.df$CumRain[is.na(IA.df$CumRain)] <- 0

rm(IA2016.df, IAprecip.df)
```

#Read in SEV
```{r readSEV}
SEV.df <- read.csv('data/SEV/sev1_meteorology_2011-2015.csv') %>%
  filter(StationID == 40) %>%
  select(Year, Julian_Day, Hour, Temp_C, Relative_Humidity, Precipitation, Wind_Dir, Mean_WindSpeed) %>%
  mutate(DateStr = sprintf('%.4d-%.3d %.2d', Year, Julian_Day, Hour)) %>%
  mutate(Date = as.POSIXct(DateStr, format='%Y-%j %H', tz='GMT'),
         Location = 'SEV') %>%
  dplyr::rename(Air.temperature=Temp_C, Relative.Humidity=Relative_Humidity, Rain=Precipitation,
         Wind.speed=Mean_WindSpeed, Wind.direction=Wind_Dir) %>%
  arrange(Date) %>%
  mutate(duration.hr = c(NA, as.numeric(Date[-1] - Date[-length(Date)], units='hours')))

SEV.df[SEV.df == -999] <- NA
SEV.df$tempRain <- SEV.df$Rain
SEV.df$tempRain[is.na(SEV.df$tempRain)] <- 0
SEV.df$CumRain <- cumsum(SEV.df$tempRain)
SEV.df$tempRain <- NULL
```

#Merge data
```{r mergeData}
data.df <- bind_rows(FED.df, GBB.df, IA.df, SEV.df) %>%
  select(Location, Date, duration.hr, Wind.direction, Wind.speed, Air.temperature, Relative.Humidity, 
         CumRain, Rain, Juvik, Wetness.perc, Leaf.wetness) %>%
  arrange(Location, Date) %>%
  mutate(Location = as.factor(Location))

data.df[data.df == -999 | data.df == -888.88] <- NA
```

#Check variables
```{r varCheck}

plot.df <- data.df %>%
  select(-Date, -CumRain, -duration.hr, -Rain) %>%
  tidyr::gather(variable, value, -Location,na.rm=TRUE)

ggplot(plot.df) + geom_histogram(aes(x=value)) +
  facet_grid(Location~variable, scales='free')

data.df %>%
  group_by(Location) %>%
  summarize(startTime=min(Date),
         endTime = max(Date),
         totalRain = max(CumRain),
         avgDuration=mean(duration.hr, na.rm=TRUE),
         totalIrregularTime.hr=sum(duration.hr[duration.hr != 1], na.rm=TRUE)) %>% print
```

#Plot noon wetness on days without rain
```{r noon}

noon.df <- data.df %>%
  filter(Air.temperature > -100 & Relative.Humidity >0) %>% 
  mutate(shortDay = as.POSIXct(trunc(Date, units='days')),
        shortHour = hour(Date)) %>%
  group_by(Location, shortDay) %>%
  summarize(dailyRain = sum(Rain, na.rm=TRUE),
            Wetness.perc = Wetness.perc[which.min(abs(shortHour - 12))],
            Leaf.wetness = Leaf.wetness[which.min(abs(shortHour - 12))],
            Relative.Humidity = Relative.Humidity[which.min(abs(shortHour - 12))],
            Air.temperature = Air.temperature[which.min(abs(shortHour - 12))]) %>%
  filter(dailyRain == 0)

ggplot(subset(noon.df, Location == 'GBB')) + 
  geom_point(aes(x=Leaf.wetness, y=Relative.Humidity, color=Air.temperature)) + 
  facet_wrap(~Location)
ggsave('temp/LeafWetness_NoonWet.pdf')

ggplot(subset(noon.df, Location %in% c('FED', 'IA'))) + 
  geom_point(aes(x=Wetness.perc, y=Relative.Humidity, color=Air.temperature)) + 
  scale_x_log10() +
  facet_wrap(~Location)
ggsave('temp/WetnessPerc_NoonWet.pdf')


```

#Set wet factor
```{r setWet}

#If the Leaf.wetness is below (yes below) 150, or above Wetness.perc 10% then it is wet 
data.df$Wet.factor <- with(data.df, 
                           factor(ifelse(Location == 'FED', #use wetness.perc
                                         ifelse(is.na(Wetness.perc), NA, 
                                                ifelse(Wetness.perc > 10, 'wet',
                                                       ifelse(Wetness.perc > 2.5, 'slightly wet', 'dry'))),
                                         ifelse(Location == 'GBB', #use Leaf.wetness
                                                ifelse(Leaf.wetness == 7999, 'dry',
                                                       ifelse(Leaf.wetness < 150, 'wet', 'slightly wet')),
                                                ifelse(Location == 'IA',
                                                       ifelse(is.na(Wetness.perc), NA, 
                                                              ifelse(Wetness.perc > 25, 'wet',
                                                                     #ifelse(Wetness.perc > 12, 'slightly wet',
                                                                     'dry')),
                                                NA))),
             levels=c('dry', 'slightly wet', 'wet')))

ggplot(subset(data.df, Location %in% c('FED', 'IA'))) + 
  geom_histogram(aes(x=Wetness.perc, fill=Wet.factor, y=..density..), alpha=0.5) +
  facet_grid(~Location)
ggplot(subset(data.df, Location %in% c('GBB'))) + 
  geom_histogram(aes(x=Leaf.wetness, fill=Wet.factor, y=..density..), alpha=0.5) +
  scale_x_log10() +
  facet_grid(~Location)

```

# Define events
## Define event start by 2hr >90% and end by <90% relative humidity
```{r countEventsByRHThreshold}
RH_threshold <- 90

data.df <- data.df %>%
  arrange(Location, Date) %>%
  group_by(Location) %>%
  mutate(wetRH=is.finite(Relative.Humidity) & Relative.Humidity >= RH_threshold,
         dryRH=is.finite(Relative.Humidity) & Relative.Humidity < RH_threshold) %>%
  mutate(RHtwoHrStart = wetRH & #currently above 90
           c(FALSE, dryRH[-length(wetRH)]) & #previously not above 90
           c(wetRH[-1], FALSE), #next is above 90
         RHdrop = dryRH & #currently dry
           c(FALSE, wetRH[-length(wetRH)]) & 
           c(FALSE, FALSE, wetRH[-length(wetRH)+0:1]) #previously wet for 2 hrs
         ) %>% 
  mutate(EventIndex.byRHThreshold = cumsum(RHtwoHrStart | RHdrop)) %>%
  ungroup() %>%
  select(-wetRH, -dryRH, -RHtwoHrStart, -RHdrop)
  
ggplot(data.df) + 
  geom_line(aes(x=Date, y=EventIndex.byRHThreshold)) + 
  facet_wrap(~Location, ncol=1, scales='free_y')

ggplot(data.df[500+10:50,], aes(x=Date, y=Relative.Humidity)) + 
  geom_line(aes(color=as.factor(EventIndex.byRHThreshold))) +
  geom_point()+
  geom_hline(yintercept=RH_threshold) +
  facet_wrap(~Location, ncol=1, scales='free_y')

```
## Define events by recorded wet factor and Rain
```{r countEventsByWet_Rain}
data.df <- data.df %>%
  group_by(Location) %>%
  mutate(wetStart.Rain = is.finite(Rain) & Rain > 0 & #currently raining
           c(TRUE, (is.na(Rain) | Rain == 0)[-length(Rain)]), #previously not raining
         dryEnd.Rain = (is.na(Rain) | Rain == 0) & #currently not raining
           c(TRUE, (is.finite(Rain) & Rain > 0)[-length(Rain)]), #previously raining
         wetStart.wf = grepl('wet', as.character(Wet.factor)) & #currently wet
           c(TRUE, grepl('dry', as.character(Wet.factor[-length(Wet.factor)]))), #previously dry
         dryEnd.wf= grepl('dry', as.character(Wet.factor)) & #currently dry
           c(TRUE, grepl('wet', as.character(Wet.factor[-length(Wet.factor)])))) %>% #previously wet) %>%
  mutate(EventIndex.byRain = cumsum(wetStart.Rain | dryEnd.Rain),
         EventIndex.byWetFactor = cumsum(wetStart.wf | dryEnd.wf)) %>%
  ungroup() %>%
  select(-wetStart.Rain, -dryEnd.Rain, -wetStart.wf, -dryEnd.wf)

data.df[data.df$Location == 'SEV', 'EventIndex.byWetFactor'] <- NA

ggplot(data.df) + 
  geom_line(aes(x=Date, y=EventIndex.byWetFactor)) + 
  facet_wrap(~Location, ncol=1, scales='free_y')

ggplot(data.df) + 
  geom_line(aes(x=Date, y=EventIndex.byRain)) + 
  facet_wrap(~Location, ncol=1, scales='free_y')
```

##Define events by wetness slope

```{r countEventsByWetSlope}
data.df <- data.df %>%
  arrange(Location, Date) %>%
  group_by(Location) %>%
  mutate(wet.slope = ifelse(is.finite(Wetness.perc), 
                            defineSlope(x=signif(Wetness.perc, digits=2)),
                            ifelse(is.finite(Leaf.wetness), 
                                   defineSlope(x = signif(-1*Leaf.wetness, digits=2)), NA))) %>%
  mutate(EventIndex.byWetSlope = defineEventIndex(wet.slope))

data.df[data.df$Location == 'SEV', 'EventIndex.byWetSlope'] <- NA
```
```{r countEventsByWetSlope_plots}
#450:500
ggplot(data.df[500-50:-50,], aes(x=Date, y=Wetness.perc,
                                color=as.factor(EventIndex.byWetSlope))) + 
  geom_point(aes(shape=wet.slope)) +geom_line()

ggplot(data.df[500+10:50,], aes(x=Date, y=Wetness.perc,
                                color=as.factor(EventIndex.byWetSlope))) + 
  geom_point(aes(shape=wet.slope)) +geom_line()


ggplot(data.df[500+10:50,], aes(x=Date, y=Relative.Humidity)) + 
  geom_line(aes(color=as.factor(EventIndex.byWetSlope))) +
  geom_point()+
  facet_wrap(~Location, ncol=1, scales='free_y')

ggplot(data.df[is.finite(data.df$Wetness.perc),], aes(x=Relative.Humidity, y=Wetness.perc)) + 
  geom_point()+
  facet_wrap(~Location, ncol=1, scales='free_y')


ggplot(data.df[is.finite(data.df$Leaf.wetness),], aes(x=Relative.Humidity, y=-1*Leaf.wetness)) + 
  geom_point(alpha=0.05)+
  facet_wrap(~Location, ncol=1, scales='free_y')

ggplot(subset(data.df, Location == 'GBB')[200:250,]) + 
  geom_line(aes(x=Date, y=-1*Leaf.wetness,
                color=as.factor(EventIndex.byWetSlope))) + 
  geom_point(aes(x=Date, y=-1*Leaf.wetness,
                color=as.factor(EventIndex.byWetSlope), shape=wet.slope)) + 
  geom_hline(yintercept=-150)
  #scale_y_log10()

ggplot(subset(data.df, Location != 'SEV')) + geom_line(aes(x=Date, y=EventIndex.byWetSlope)) + facet_wrap(~Location, ncol=1, scales='free_y')
```

##ID event types by different indecies
```{r compareIndex}
plot.df <- data.df %>% select(Location, contains('EventIndex'))

GGally::ggpairs(plot.df, columns=2:5, ggplot2::aes(color=Location, alpha=0.1))
```

# Tally lengths by event types
```{r characterizeEvents}

events.df <- data.df %>% 
  select(Date, Location, duration.hr, Wet.factor, Rain, Juvik, Relative.Humidity, Air.temperature,
         EventIndex.byRain, EventIndex.byWetFactor, EventIndex.byRHThreshold, EventIndex.byWetSlope) %>% 
  dplyr::rename(byRain=EventIndex.byRain, WetFactor=EventIndex.byWetFactor,
                RHThreshold=EventIndex.byRHThreshold, WetSlope=EventIndex.byWetSlope) %>%
  tidyr::gather(EventDefined, index, byRain, WetFactor, RHThreshold, WetSlope) %>%
  group_by(Location, EventDefined, index) %>%
  dplyr::summarise(
            EventDuration.hr=sum(duration.hr, na.rm=TRUE), 
            DryCount = sum(as.character(Wet.factor) %in% 'dry', na.rm=TRUE),
            FiniteWFCount = sum(is.finite(Wet.factor)),
            Rain=sum(Rain, na.rm=TRUE),
            RainStatus=ifelse(any(is.finite(Rain) & Rain > 0), 'Rain', 'NoRain'),
            FogStatus=ifelse(all(is.finite(Juvik)), ifelse(any(Juvik > 0), 'Fog', 'NoFog'), 'NA'),
          startTime=min(Date), endTime=max(Date),
          avgRH = mean(Relative.Humidity, na.rm=TRUE),
          avgT = mean(Air.temperature, na.rm=TRUE)) %>%
  mutate(DryPercent=ifelse(FiniteWFCount==0, NA, DryCount/FiniteWFCount)) %>%
  filter(is.finite(index))

#checkHrs <- data.df %>% group_by(Location) %>% summarize(time.hr=sum(duration.hr, na.rm=TRUE))
#checkHrs2 <- events.df %>% group_by(Location, EventDefined)  %>% summarize(time.hr=sum(EventDuration.hr, na.rm=TRUE))

```

## How many hours wet
```{r countHours}

events.df <- events.df %>%
  mutate(Type = as.factor(ifelse(EventDefined == 'byRain', 
                       ifelse(RainStatus == 'Rain', 'Wet', 'Dry'), #define rain events by rain presence
                       ifelse(EventDefined == 'WetFactor', #define wet.factor by wet.factor status
                              ifelse(is.na(DryPercent),NA, ifelse(DryPercent ==  1, 'Dry', 'Wet')),
                              ifelse(EventDefined == 'RHThreshold',  #define high RH by averate RH
                                     ifelse(round(avgRH) >= 90, 'Wet', 'Dry'),
                                     ifelse(EventDefined == 'WetSlope', #define slope by % wet.factor
                                            ifelse(is.na(DryPercent),
                                                   NA,  
                                                   ifelse(DryPercent > 0.8, 'Dry', 'Wet'))))))))  %>%
  #assign rain vs NRM event triggers
  mutate(Trigger = ifelse(Type == 'Dry', 'NA', ifelse(RainStatus == 'NoRain', 'NRM', 'Rain'))) 

wetHours <- events.df %>% 
  filter(Type == 'Wet') %>%
  group_by(Location, EventDefined, Trigger) %>%
  summarize(duration.hr=sum(EventDuration.hr)) %>%
  tidyr::spread(EventDefined, duration.hr)

kable(wetHours)

```
  
  
```{r WetEventDurations}
ggplot(subset(events.df, Type=='Wet'), 
       aes(x=EventDuration.hr, fill=Trigger))+
  geom_density(aes(y=..count..), alpha=0.5) +
         facet_grid(EventDefined~Location, scales='free') +
  scale_x_log10()+
  labs(title='Duration of wet events')

#####
#SE graphing duration histogram figure for Manuscript with sites combined, and with one definition (Wet Slope?)
ggplot(subset(events.df, Type=='Wet' 
              & EventDefined=="RHThreshold"),  #Thought we should pick one to go forward with- I picked WetSlope
       aes(x=EventDuration.hr, fill=Trigger))+
  #geom_density(aes(y=..count..), alpha=0.5) +
  geom_density(alpha = 0.2)+
  facet_grid(~Location, scales='free') +
  #scale_x_log10()+
  labs(title='Duration of wet events', x='Duration (hour)') +
  xlim(c(0,24))

# Try plotting with RH threshold definition, looking at FED and GBB, there is something wrong. I cannot seem to identify the problematic datapoints..

#this does not look right! Rain events are rarely this long except maybe in Iowa. Also, did we not exclude "events" less than 2 hours?
##KTB: While it might not rain for more then 2hours the moisture event associated with that rainfall is longer according to this analysis so it's listed as a 'rain' event according to this definition.

#Temperature histogram
ggplot(events.df %>% filter(Type=='Wet', EventDefined=="WetSlope"), 
       aes(x=avgT, fill=Trigger)) +
  geom_density(alpha=0.2) +
  facet_grid(~Location, scales='free') +
  #scale_x_log10()+
  labs(title='Average temperature of wet events', x='Air Temperature (C)') 


#Numbers estimates
events.df %>% 
  filter(Type=='Wet', EventDefined=="WetSlope") %>%
  group_by(Location, Trigger) %>%
  summarize(meanT=mean(avgT, na.rm=TRUE), sdT=sd(avgT, na.rm=TRUE),
            meanDuration=mean(EventDuration.hr, na.rm=TRUE), sdDuration=sd(EventDuration.hr, na.rm=TRUE))
```

# Relative humidity frequency by event type
```{r RHFreqByEvent}

RHbyEvents.df <- data.df %>% 
  select(Date, Location, duration.hr, Wet.factor, Rain, Juvik, Relative.Humidity,
         EventIndex.byRain, EventIndex.byWetFactor, EventIndex.byRHThreshold, EventIndex.byWetSlope) %>% 
  dplyr::rename(byRain=EventIndex.byRain, WetFactor=EventIndex.byWetFactor,
                RHThreshold=EventIndex.byRHThreshold, WetSlope=EventIndex.byWetSlope) %>%
  tidyr::gather(EventDefined, index, byRain, WetFactor, RHThreshold, WetSlope) %>%
  merge(events.df[,c('Location', 'EventDefined', 'index', 'DryPercent', 'Type', 'Trigger')], 
        all=TRUE, by=c('Location', 'EventDefined', 'index')) %>%
  rename(Event.DryPercent=DryPercent, Event.Type=Type, Event.Trigger=Trigger) %>%
  arrange(Location, EventDefined, Date)

ggplot(RHbyEvents.df, aes(x=Relative.Humidity, fill=Event.Trigger)) + 
  geom_density(alpha=0.5) +
  coord_cartesian(ylim=c(0, 0.1)) +
  facet_grid(Location ~ EventDefined)

```

##Generate probabilities of wet/dry based on RH without Rain
```{r probRH}

RHcuts <- seq(0, 100, by=4)
RHprob.df <- RHbyEvents.df %>%
  filter(Event.Trigger %in% c('NA', 'NRM'), Location != 'SEV') %>% 
  mutate(Relative.Humidity_bin=cut(Relative.Humidity, breaks=RHcuts, labels=FALSE),
         Wet.factor_short=ifelse(grepl('wet', Wet.factor), 'Wet', 
                                 ifelse(grepl('dry', Wet.factor), 'Dry', 'NA'))) %>%
  group_by(Location, EventDefined, Wet.factor_short, Relative.Humidity_bin) %>%
  tally %>%
  filter(n > 10) %>%
  ungroup() %>% group_by(Location, EventDefined, Relative.Humidity_bin) %>%
  mutate(percentTot = n/sum(n),
         min_RH = RHcuts[Relative.Humidity_bin], 
         max_RH = RHcuts[Relative.Humidity_bin+1]) %>% 
  mutate(mean_RH=(min_RH+max_RH)/2)

#To include in supplemental - showing robustness across sites and definitions
ggplot(subset(RHprob.df, Relative.Humidity_bin > (length(RHcuts)/2))) +
  geom_line(aes(x=mean_RH, y=percentTot, color=Wet.factor_short)) +
  facet_grid(Location~EventDefined, scales='free') +
  labs(x='Relative humidity (%)', 
       y='Leaf wetness likelihood (%)', colour="Wet/Dry factor")
#Is y-axis a likelihood or an actual wetness?

thresholds.df <- RHprob.df %>%
  filter(Wet.factor_short == 'Wet') %>%
  ungroup() %>% group_by(Location, EventDefined) %>%
  summarize(bin_RH = Relative.Humidity_bin[which.min(abs(percentTot-0.5))]) %>%
  ungroup() %>%
  mutate(min_RH = RHcuts[bin_RH], max_RH = RHcuts[bin_RH+1]) %>%
  mutate(mean_RH = (min_RH+max_RH)/2)

kable(thresholds.df)

likelihood.df <- RHprob.df %>%
  filter(Wet.factor_short == 'Wet', is.finite(Relative.Humidity_bin)) %>%
  ungroup() %>% group_by(Relative.Humidity_bin) %>%
  summarize(percentWet = mean(percentTot)) %>%
  mutate(min_RH = RHcuts[Relative.Humidity_bin], 
         max_RH = RHcuts[Relative.Humidity_bin+1])%>% 
  mutate(mean_RH=(min_RH+max_RH)/2)

likelihood.df <- rbind(likelihood.df,
                       data.frame(Relative.Humidity_bin=1:min(likelihood.df$Relative.Humidity_bin),
                                  percentWet=0, 
                                  min_RH=0, 
                                  max_RH=min(likelihood.df$min_RH),
                                  mean_RH=min(likelihood.df$min_RH)/2))

ggplot(likelihood.df, aes(x=mean_RH, y=percentWet)) +
  geom_line() + geom_point() +
  labs(x='Relative humidity', y='Percent wet (with removed rain events)', 
       title='Likelihood wet across all locations and event definitions')
```

##Compare wet times
```{r compareWet}
data_addWetL.df <- data.df %>%
  mutate(Relative.Humidity_bin = cut(Relative.Humidity, breaks=RHcuts, labels=FALSE)) %>%
  merge(likelihood.df[,c('Relative.Humidity_bin', 'percentWet')]) %>%
  rename(percentWet_byRH=percentWet) %>%
  select(-Relative.Humidity_bin) %>%
  arrange(Location, Date)

compareTime.df <- data_addWetL.df %>%
  ungroup() %>% group_by(Location) %>%
  summarize(total.hr=sum(duration.hr, na.rm=TRUE),
         rain.hr=sum(duration.hr*(Rain > 0), na.rm=TRUE),
         highRH.hr=sum(duration.hr*(Relative.Humidity>80), na.rm=TRUE),
         wet.hr=sum(duration.hr*(grepl('wet', Wet.factor)), na.rm=TRUE),
         wet_by_RH.hr=sum(duration.hr*percentWet_byRH, na.rm=TRUE))

kable(compareTime.df)
```

#Duration
```{r}
dayData.df <- data.df %>%
  mutate(Day = yday(Date), Year=year(Date), agg='Day') %>%
  group_by(Location, Day, Year, agg) %>%
  summarize(duration.day=sum(duration.hr, na.rm=TRUE)/24, 
            mean.temperature=mean(Air.temperature, na.rm=TRUE), 
            min.temeprature=min(Air.temperature, na.rm=TRUE),
            max.temeprature=min(Air.temperature, na.rm=TRUE),
            mean.RH=mean(Relative.Humidity, na.rm=TRUE),
            min.RH=min(Relative.Humidity, na.rm=TRUE),
            max.RH=max(Relative.Humidity, na.rm=TRUE),
            CumRain=max(CumRain, na.rm=TRUE),
            Rain=sum(Rain, na.rm=TRUE),
            Juvik.hr=sum(duration.hr*(Juvik > 0)),
            Wet.hr=ifelse(all(is.na(Wet.factor)), 
                          NA, sum(duration.hr*(grepl('wet', Wet.factor)), na.rm=TRUE)))

weekData.df <- data.df %>%
  mutate(Week = week(Date), Year=year(Date), agg='Week') %>%
  group_by(Location, Week, Year, agg) %>%
  summarize(duration.day=sum(duration.hr, na.rm=TRUE)/24, 
            mean.temperature=mean(Air.temperature, na.rm=TRUE), 
            min.temeprature=min(Air.temperature, na.rm=TRUE),
            max.temeprature=min(Air.temperature, na.rm=TRUE),
            mean.RH=mean(Relative.Humidity, na.rm=TRUE),
            min.RH=min(Relative.Humidity, na.rm=TRUE),
            max.RH=max(Relative.Humidity, na.rm=TRUE),
            CumRain=max(CumRain, na.rm=TRUE),
            Rain=sum(Rain, na.rm=TRUE),
            Juvik.hr=sum(duration.hr*(Juvik > 0)),
            Wet.hr=ifelse(all(is.na(Wet.factor)), 
                          NA, sum(duration.hr*(grepl('wet', Wet.factor)), na.rm=TRUE)))

monthData.df <- data.df %>%
  mutate(Month = month(Date), Year=year(Date),agg='Month') %>%
  group_by(Location, Month, Year, agg) %>%
  summarize(duration.day=sum(duration.hr, na.rm=TRUE)/24, 
            mean.temperature=mean(Air.temperature, na.rm=TRUE), 
            min.temeprature=min(Air.temperature, na.rm=TRUE),
            max.temeprature=min(Air.temperature, na.rm=TRUE),
            mean.RH=mean(Relative.Humidity, na.rm=TRUE),
            min.RH=min(Relative.Humidity, na.rm=TRUE),
            max.RH=max(Relative.Humidity, na.rm=TRUE),
            CumRain=max(CumRain, na.rm=TRUE),
            Rain=sum(Rain, na.rm=TRUE),
            Juvik.hr=sum(duration.hr*(Juvik > 0)),
            Wet.hr=ifelse(all(is.na(Wet.factor)), 
                          NA, sum(duration.hr*(grepl('wet', Wet.factor)), na.rm=TRUE)))

yearData.df <- data.df %>%
  mutate(Year=year(Date), agg='Year') %>%
  group_by(Location, Year, agg) %>%
  summarize(duration.day=sum(duration.hr, na.rm=TRUE)/24, 
            mean.temperature=mean(Air.temperature, na.rm=TRUE), 
            min.temeprature=min(Air.temperature, na.rm=TRUE),
            max.temeprature=min(Air.temperature, na.rm=TRUE),
            mean.RH=mean(Relative.Humidity, na.rm=TRUE),
            min.RH=min(Relative.Humidity, na.rm=TRUE),
            max.RH=max(Relative.Humidity, na.rm=TRUE),
            CumRain=max(CumRain, na.rm=TRUE),
            Rain=sum(Rain, na.rm=TRUE),
            Juvik.hr=sum(duration.hr*(Juvik > 0)),
            Wet.hr=ifelse(all(is.na(Wet.factor)), 
                          NA, sum(duration.hr*(grepl('wet', Wet.factor)), na.rm=TRUE)))

aggData <- data.df %>% 
  mutate(agg='Hour', Wet.hr=ifelse(grepl('wet', Wet.factor), duration.hr, 0), 
         Juvik.hr=ifelse(is.na(Juvik), NA, ifelse(Juvik>0, duration.hr, 0)),
         duration.day=duration.hr/24, Rain=ifelse(is.na(Rain), 0, Rain)) %>%
  rename(mean.RH=Relative.Humidity, mean.temperature=Air.temperature, 
         Hour=Date) %>%
  select(Location, agg, Wet.hr, Juvik.hr, duration.day, mean.RH, mean.temperature, Hour, CumRain, Rain) %>%
  bind_rows(dayData.df, weekData.df, monthData.df, yearData.df) %>%
  mutate(agg=factor(agg, levels=c('Hour', 'Day', 'Week', 'Month', 'Year')),
         hasRain = ifelse(Rain > 0, 'Rain', 'NoRain')) 
```

```{r timeComp}
ggplot(subset(data.df, grepl('wet', Wet.factor))) +
  geom_density(aes(x=Relative.Humidity, color=Location)) +
  labs(title='Hour wet sensor is active')

ggplot(subset(aggData, Location != 'SEV')) +
  geom_jitter(aes(y=Wet.hr/duration.day, x=mean.RH, color=Location)) +
  facet_grid(hasRain~agg)#, ncol=1)
  
triggerHours <- aggData %>%
  filter(Location != 'SEV') %>%
  group_by(Location, agg, hasRain) %>%
  summarize(Wet.hr = sum(Wet.hr, na.rm=TRUE), duration.day=sum(duration.day, na.rm=TRUE)) %>% 
  group_by(Location, agg) %>%
  mutate(totalWet.hr = sum(Wet.hr))

trueWet <- data_addWetL.df %>%
  group_by(Location, EventIndex.byWetSlope) %>%
  summarize(tempWet.hr=sum(duration.hr[grepl('wet', Wet.factor)], na.rm=TRUE), 
         hasRain=ifelse(any(is.finite(Rain) & Rain>0), 'Rain', 'NoRain')) %>%
  group_by(Location, hasRain) %>%
  summarize(Wet.hr=sum(tempWet.hr)) %>%
  group_by(Location) %>%
  mutate(totalWet.hr = sum(Wet.hr))

ggplot(triggerHours, aes(x=agg, y=Wet.hr, color=as.factor(Location), group=Location)) +
  geom_line() +geom_point() + 
  geom_hline(data=trueWet, aes(yintercept=Wet.hr, color=as.factor(Location)), linetype=2) +
  scale_y_log10() + 
  facet_wrap(~hasRain) +
  labs(x='Temporal aggregation level', y='Hours wet', 
       title='Hours wet attributed to rain events')

ggplot(triggerHours, aes(x=agg, y=Wet.hr/totalWet.hr, color=as.factor(Location), group=Location)) +
  geom_line() +geom_point() + 
  geom_hline(data=trueWet, aes(yintercept=Wet.hr/totalWet.hr, color=as.factor(Location)), linetype=2) +
  scale_y_log10() + 
  facet_wrap(~hasRain) +
  labs(x='Temporal aggregation level', y='Fraction of wet hours', 
       title='Hours wet attributed to rain events')
```
#Save data
```{r saveData}
save(data_addWetL.df, compareTime.df, file='temp/MetData20170903.RData')
```

#Helper functions
```{r, comment=''}
for(xx in sourceFiles.ls){
  print(readLines(xx))
}

```